Star-galaxy Classification Using Deep Convolutional Neural Networks

### 2016-06-18 paper/main.tex
added robert's comments

@@

\begin{abstract}
Most existing star-galaxy classifiers use the reduced summary information from catalogs,
requiring careful feature extraction and selection.
The latest advances in machine learning using deep convolutional neural networks
allow a machine to automatically learn the features directly from data,
minimizing need for input from human experts.
We present a star-galaxy classification framework that uses
deep convolutional neural networks (ConvNets) directly on the raw pixel values.
Using data from the Sloan Digital Sky Survey (SDSS) and
the Canada-France-Hawaii Telescope Lensing Survey (CFHTLenS),
we demonstrate that ConvNets are able to produce accurate and well-calibrated
probabilistic classifications that are competitive with
conventional machine learning techniques.
Future advances in deep learning may bring more success in currently ongoing and
forthcoming photometric surveys, such as the Dark Energy Survey (DES) and
the Large Synoptic Survey Telescope (LSST), because deep neural networks require
very little feature engineering by hand.
\end{abstract}

@@

Currently ongoing and forthcoming large-scale photometric surveys,
such as the Dark Energy Survey (DES) and the Large Synoptic Survey Telescope (LSST),
aim to collect photometric data for hundreds of millions to billions of
stars and galaxies.
Due to the sheer volume of data, it is not possible for human experts to
manually classify them,
and the separation of photometric catalogs into stars and galaxies has to be automated.
Furthermore, any classification approach must be probabilistic in nature.
A fully probabilistic classifier enables a user to adopt probability cuts
to obtain a pure sample for population studies,
or to optimize the allocation of observing time by selecting objects for follow-up.
However, ideally, the probability estimates themselves would be retained for all sources
and used in subsequent analyses to improve or enhance a particular
measurement~\citep{ross2011ameliorating,seo2012acoustic}.

With machine learning, we can use algorithms to automatically create
accurate source catalogs with well-calibrated posterior probabilities.
Machine learning techniques have been a popular tool
in many areas of astronomy~\citep{ball2007robust,ball2008robust,
banerji2010galaxy,
carrascokind2013tpz,carrascokind2014somz,carrascokind2014exhausting,
ivezic2014statistics,
kamdar2016machine1,kamdar2016machine2}, and
artificial neural networks were first applied to the problem of star-galaxy
classification in the work of \citet{odewahn1992automated},
which has become a core part of the astronomical image processing software
\texttt{SExtractor}~\citep{bertin1996sextractor}.
Other successfully implemented examples of applying machine learning to
the star-galaxy classification problem include decision 
trees~\citep{weir1995automated,suchkov2005census,ball2006robust,sevilla2015effect}
, Support Vector Machines~\citep{Fadely2012}, and
classifier combination strategies~\citep{kim2015hybrid}.

Almost all star-galaxy classifiers published in the literature use
the reduced summary information available from astronomical catalogs.
Constructing catalogs requires careful engineering and considerable domain expertise
to transform the raw pixel values of an image to suitable features,
such as magnitudes or shape information of an object.
In a branch of machine learning called \textit{deep learning},
features are not designed by human experts;
they are learned directly from data by deep neural networks.
Deep learning has been applied successfully to galaxy morphological
classification in Sloan Digital Sky Survey~\citep[SDSS;][]{dieleman2015rotation}
and Cosmic Assembly Near-infrared Deep Extragalactic Legacy
Survey~\citep[CANDELS;][]{huertas2015catalog} and to photometric redshift
estimation~\citep{hoyle2015measuring},
but it has never been applied to the problem of source classification.

In this paper, we present a star-galaxy classification framework
that uses a convolutional neural network (ConvNet) model directly on the images
from SDSS and Canada-France-Hawaii Telescope Lensing Survey (CFHTLenS).
We compare its performance with a standard machine learning technique
that uses the reduced summary information from catalogs,
and demonstrate that our ConvNet model is able to produce accurate and
well-calibrated probabilistic classifications with very little feature
engineering by hand.
In Section~\ref{sec:data}, we describe the data sets used in this paper
and the pre-processing steps for preparing the image data sets.
We provide a brief overview of deep learning and ConvNets
in Section~\ref{sec:deep_learning}, and
discuss our strategy for preventing overfitting in Section~\ref{sec:overfitting}.
In Section~\ref{sec:tpc}, we describe a state-of-the-art tree-based
machine learning algorithm, to which the performance of our ConvNet model is
compared.
We present the main results of our ConvNet model in
Section~\ref{sec:results_and_discussion}, and
we outline our conclusions in Section~\ref{sec:conclusions}.

@@

\begin{figure}
  \centering
  \includegraphics[width=\columnwidth]{figures/sdss_color_color.pdf}
  \caption{
    Hexbin plots showing the color-color space distribution of stars and
    galaxies in the SDSS data set.
    It is easy to see that there is significant overlap between stars
    and galaxies in the color-color space. 
  }
  \label{fig:sdss_color_color}
\end{figure}

The Sloan Digital Sky Survey~\citep[SDSS;][]{york2000sloan}
phases I--III obtained photometric data in five bands,
$u$, $g$, $r$, $i$, and $z$,
covering 14,555 square degrees, more than one-third of the entire sky.
The resulting catalog contains photometry of over 300 million stars galaxies
with a limiting magnitude of $r \approx 22$,
making the SDSS one of the largest sky surveys ever undertaken.
The SDSS also conducted an expansive spectroscopic follow-up of
more than three million stars and galaxies~\citep{eisenstein2011sdss}.
In this paper, we use a subset of the photometric and spectroscopic data
contained within the Data Release 12~\citep[DR12;][]{alam2015eleventh},
which is publicly available through the online CasJobs
server\footnote{http://skyserver.sdss.org/casjobs/}~\citep{li2008casjobs}.

Using the CasJobs server, we randomly select $65\,000$ stars and galaxies.
We exclude some bad photometric observations as follows.
We consider only objects whose $r$-band magnitude is less than 40;
there are no warning flags in the spectroscopic measurement (\texttt{zWarning = 0});
the half-light radius in the $r$ band is less than 30 arc seconds
as measured by the exponential and de Vaucouleurs light profiles;
the error on the spectroscopic redshift measurement is less than 0.1; and
the spectroscopic redshift is less than 2.

To create training images,
we obtain the image FITS files for SDSS fields containing these objects
in five photometric bands: $u$, $g$, $r$, $i$, and $z$.
We use the astrometry information in the FITS headers
in the \texttt{Montage}\footnote{http://montage.ipac.caltech.edu/} software
to align each image to the reference ($r$-band) image.
We then use \texttt{SExtractor} to find the pixel positions of
the $65\,000$ objects we have selected,
and to center each object in the cutout image.
We also convert the flux values to inverse hyperbolic sine magnitudes
\citep[also known as luptitudes;][]{lupton1999modified}.
Finally, in order to account for the effect of galactic dust,
galactic extinction corrections in magnitudes are applied
following \cite{schlegel1998maps}.
In the end, we have cutout images of size $48\times48$ pixels
with luptitude values in each pixel.
We note that we have experimented with increasing the pixel dimensions to
$60\times60$ and $72\times72$ pixels, but do not find noticeable
improvement in the performance of our model.

In the end, we have $17\,344$ stars and $47\,656$ galaxies available
for the training and testing processes.
The apparent magnitudes range from $10.7 < r < 23.1$,
and the galaxies in this sample have a mean redshift of $z \sim 0.36$.
The $ugriz$ color-color space of the sample is plotted in
Figure~\ref{fig:sdss_color_color}.
We randomly split the objects into training, held-out validation,
and blind test sets of size $40\,000$, $10\,000$, and $15\,000$, respectively.
We note that cross-validation is often avoided in deep learning
in favor of hold-out validation,
since cross-validation is computationally expensive.
We also note that we perform a blind test, and the test set is not used in any
way to train or calibrate the algorithms.

@@

\begin{figure}
  \centering
  \includegraphics[width=\columnwidth]{figures/clens_color_color.pdf}
  \caption{
    Hexbin plots showing the color-color space distribution of stars and
    galaxies in the CFHTLenS data set.
  }
  \label{fig:clens_color_color}
\end{figure}

We also use photometric data from
the Canada-France-Hawaii Telescope Lensing Survey
\cite[CFHTLenS\footnote{http://www.cfhtlens.org/};][]
{heymans2012cfhtlens,erben2013cfhtlens,hildebrandt2012cfhtlens}.
This catalog consists of more than twenty five million objects
with a limiting magnitude of $i_{\text{AB}} \approx 25.5$. 
It covers a total of 154 square degrees
in the four fields (named W1, W2, W3, and W4)
of CFHT Legacy Survey~\citep[CFHTLS;][]{gwyn2012canada}
observed in the five photometric bands:
$u$, $g$, $r$, $i$, and $z$.

@@

We obtain FITS images for each 1 square degree CFHTLenS pointing
that contains objects with spectroscopic labels.
We create cutout images of size $96\times96$ pixels by using a similar method
to that described in Section~\ref{sec:sdss}.
Finally, images are downscaled to $48\times48$ pixels
to reduce computational cost.

In the end, we have $8\,545$ stars and $57\,843$ galaxies available
for the training and testing processes.
The apparent magnitudes range from $13.9 < r < 25.6$,
and the galaxies in this sample have a mean redshift of $z \sim 0.59$.
The $ugriz$ color-color space of the sample is plotted in
Figure~\ref{fig:clens_color_color}.
We randomly split the objects into training, held-out validation, and
blind test sets of size $40\,000$, $10\,000$, and $13\,278$, respectively.

\section{Deep Learning}
  \label{sec:deep_learning}

@@

The sigmoid function, $\sigma(x)=1/\left(1+e^{-x}\right)$,
has frequently been used as activation function, but it has recently fallen out of favor,
and the hyperbolic tangent function, $\sigma(x)=\tanh(x)$, is almost always used instead.
The most popular non-linearity at present, however, is the rectified linear unit
\citep[ReLU;][]{nair2010rectified}, $\sigma(x)=\max(0, x)$.
ReLUs generally allow much faster
training of deep neural networks with many layers.
However, ReLU units can sometimes result in dead neurons whose output is always zero.
To mitigate this problem, we use leaky ReLUs that have a small, non-zero slope
in the negative region,
\begin{equation}
  \sigma(x) =
    \begin{cases}
      x & \mbox{if } x \geq 0 \\
      0.01x & \mbox{if } x < 0.
    \end{cases}
\end{equation}

Many deep learning models use feedforward neural network architectures with multiple layers,
where each neuron in one layer is connected to the neurons of the subsequent layer.
A schematic representation is shown in Figure~\ref{fig:neuron_b}.
All layers except the input and output layers are conveniently called hidden layers.
With multiple hidden layers, deep neural networks can model complex nonlinear
relationships, and they have had great success in multimedia tasks where humans traditionally
outperformed computers: speech recognition, language processing, and image classification.

@@

@@

\begin{table*}
  \centering
  \caption{Summary of ConvNet architecture and hyperparameters. Note that pooling layers have no learnable parameters.}
  \label{table:hyperparamters}
  \begin{tabular}{ccccccc}
    \hline
    type            & filters & filter size & padding & non-linearity & initial weights & initial biases \\
    \hline
    convolutional   & 32          & $5\times5$  & -       & leaky ReLU    & orthogonal      & 0.1            \\
    convolutional   & 32          & $3\times3$  & 1       & leaky ReLU    & orthogonal      & 0.1            \\
    pooling         & -           & $2\times2$  & -       & -             & -               & -              \\
    convolutional   & 64          & $3\times3$  & 1       & leaky ReLU    & orthogonal      & 0.1            \\
    convolutional   & 64          & $3\times3$  & 1       & leaky ReLU    & orthogonal      & 0.1            \\
    convolutional   & 64          & $3\times3$  & 1       & leaky ReLU    & orthogonal      & 0.1            \\
    pooling         & -           & $2\times2$  & -       & -             & -               & -              \\
    convolutional   & 128         & $3\times3$  & 1       & leaky ReLU    & orthogonal      & 0.1            \\
    convolutional   & 128         & $3\times3$  & 1       & leaky ReLU    & orthogonal      & 0.1            \\
    convolutional   & 128         & $3\times3$  & 1       & leaky ReLU    & orthogonal      & 0.1            \\
    pooling         & -           & $2\times2$  & -       & -             & -               & -              \\
    fully-connected & 2048        & -           & -       & leaky ReLU    & orthogonal      & 0.01            \\
    fully-connected & 2048        & -           & -       & leaky ReLU    & orthogonal      & 0.01            \\
    fully-connected & 2           & -           & -       & softmax       & orthogonal      & 0.01            \\
    \hline
  \end{tabular}
\end{table*}

@@

We have performed a manual search to explore more than 200 combinations of different architectures
and hyperparameters to find an architecture that minimizes the loss function
(Equation~\ref{eq:cross_entropy_all}) on the validation set of the SDSS data.
The architecture described in this section provides the best performance on
the SDSS validation set.
To test how this model performs across different, related problems,
we use the same architecture on the CFHTLenS data set.

@@

We note that training our network takes about 40 hours on a NVIDIA Tesla K40 GPU.

@@

Our convolution neural network has $11\times10^6$ learnable parameters,
while there are only $4\times10^4$ images in the training set.
As a result, the model is very likely to \emph{overfit} without regularization.
In this section, we describe the techniques used to combat overfitting.

@@

We continue the sampling process until
we reach a predefined number of combinations,
and finally use Equation~\ref{eq:posterior_bmc} to compute
the posterior probability that a source is a star (or a galaxy).
In this paper, we use a $q$ value of three,
and $1\,000$ model combinations are considered.

@@

TPC uses classification trees,
a type of prediction trees that are designed to
provide a classification or predict a discrete category.
Prediction trees are built by asking a sequence of questions
that recursively split the data into branches
until a terminal leaf is created
that meets a stopping criterion
(\eg a minimum leaf size).
The optimal split dimension is decided by
choosing the attribute that maximizes
the \textit{Information Gain} ($I_G$), which is defined as

@@

\noindent
where $f_g$ is the fraction of galaxies in the training data.
At each node in our tree,
we scan all dimensions to identify the split point that
maximizes the information gain as defined by Equation~\ref{eq:information_gain},
and select the attribute that maximizes the impurity index overall.

@@

We train two TPC models on the SDSS data set using different sets of
attributes.
The first model, which we denote $\rmn{TPC}_{\rmn{phot}}$, is trained with
only nine photometric attributes:
the extinction-corrected model magnitudes in five bands
($u$, $g$, $r$, $i$, $z$)
and their corresponding colors
($u-g$, $g-r$, $r-i$, $i-z$).
The second model, which we denote $\rmn{TPC}_{\rmn{morph}}$, is trained with
the concentration parameter in each band in addition to the magnitudes and
colors, for a total of fifteen dimensions.
The concentration is defined as the difference between the cmodel
and PSF magnitude,
\ie $\rmn{concentration}\equiv\rmn{psfMag}-\rmn{cModelMag}$.
The SDSS pipeline uses a parametric method based on the concentration,
\ie an object is classified as a galaxy if $\rmn{concentration}>0.145$.
We find that the concentration is an excellent morphological feature
for star-galaxy separation, and including more morphological features
does not show noticeable improvement in performance.
The concentration is a good example of carefully handcrafted feature extraction,
and we show in Section~\ref{sec:results_and_discussion}
that ConvNets do not require such feature engineering.

@@

Probabilistic classification models can be considered as
functions that output a probability estimate of each source
to be in one of the classes (\eg a star or a galaxy).
The simplest way to way to use the probability output is to
convert it into a class label by using a threshold (a probability cut).
Although many studies often choose the threshold by setting it to a fixed value,
\eg $P_{\rmn{cut}} = 0.5$ \citep[\eg][]{henrion2011bayesian, Fadely2012},
choosing $0.5$ as a threshold is not the best choice
for an unbalanced data set, where galaxies outnumber stars.
Furthermore, setting a fixed threshold ignores the operating condition 
(\eg science requirements, stellar distribution, misclassification costs)
where the model will be applied.

@@

We also use the mean squared error
(MSE; also known as Brier score~\citep{brier1950verification})
as a performance metric. We define MSE as
\begin{equation} \label{eq:mse}
  \rmn{MSE} = \frac{1}{N} \sum^{N}_{j=1}
  \left( y_j - \hat{y}_j \right)^2,
\end{equation}
The MSE can be considered as both a score function that quantifies
how well a set of probabilistic predictions is calibrated,
or a loss function.

@@

The problem with binning approach is either too few or too many bins
can distort the evaluation of calibration performance.
Thus, we adopt a calibration measure based on
overlapping binning~\citep{caruana2004data}.
We order the predicted values $P_\rmn{class}$ and
put the first $1\,000$ elements in the first bin.
We calculate the true probability $P_\rmn{gal}$ by counting the true galaxies
in this bin.
The calibration error for this bin is $\mid P_\rmn{gal} - P_\rmn{class} \mid$.
We then repeat this for the second bin (2 to $1\,001$),
the third bin (3 to $1\,002$), and so on, and average the binned calibration errors.
Thus, the overall calibration error is given by
\begin{equation}
  \rmn{CAL} = \frac{1}{N-s} \sum_{b=1}^{N-s} \sum_{j=b}^{b+s-1}
  \bigg| P_{\rmn{class},j} - \frac{\sum_{j=b}^{b+s-1} P_{\rmn{gal},j}}{s} \bigg|,
\end{equation}
where $s=1\,000$ is the bin length.

>>>>

\newcommand{\RC}[1]{{\color{red}\textbf{Robert Comment: #1\ }}}

@@

\begin{abstract}
Most existing star-galaxy classifiers use the reduced summary information from catalogs,
requiring careful feature extraction and selection.
The latest advances in machine learning that use deep convolutional neural networks
allow a machine to automatically learn the features directly from data,
minimizing the need for input from human experts.
We present a star-galaxy classification framework that uses
deep convolutional neural networks (ConvNets) directly on the reduced, calibrated pixel values.
Using data from the Sloan Digital Sky Survey (SDSS) and
the Canada-France-Hawaii Telescope Lensing Survey (CFHTLenS),
we demonstrate that ConvNets are able to produce accurate and well-calibrated
probabilistic classifications that are competitive with
conventional machine learning techniques.
Future advances in deep learning may bring more success with current and
forthcoming photometric surveys, such as the Dark Energy Survey (DES) and
the Large Synoptic Survey Telescope (LSST), because deep neural networks require
very little, manual feature engineering.
\end{abstract}

@@

Currently ongoing and forthcoming large-scale photometric surveys,
such as the Dark Energy Survey (DES) and the Large Synoptic Survey Telescope (LSST),
aim to collect photometric data for hundreds of millions to billions of
stars and galaxies.
Due to the sheer volume of data, it is not possible for human experts to
manually classify them,
and the separation of photometric catalogs into stars and galaxies has to be automated.
Furthermore, any classification approach must be probabilistic in nature.
A fully probabilistic classifier enables a user to adopt probability cuts
to obtain a pure sample for population studies,
or to optimize the allocation of observing time by selecting objects for follow-up.
Ideally, however, the probability estimates themselves would be retained for all sources
and used in subsequent analyses to improve or enhance a particular
measurement~\citep{ross2011ameliorating,seo2012acoustic}.

With machine learning, we can use algorithms to automatically create
accurate source catalogs with well-calibrated posterior probabilities.
Machine learning techniques have been a popular tool
in many areas of astronomy~\citep{ball2007robust,ball2008robust,
banerji2010galaxy,
carrascokind2013tpz,carrascokind2014somz,carrascokind2014exhausting,
ivezic2014statistics,
kamdar2016machine1,kamdar2016machine2}. 
Artificial neural networks were first applied to the problem of star-galaxy
classification in the work of \citet{odewahn1992automated},
and they have become a core part of the astronomical image processing software
\texttt{SExtractor}~\citep{bertin1996sextractor}.
Other successfully implemented examples of applying machine learning to
the star-galaxy classification problem include decision 
trees~\citep{weir1995automated,suchkov2005census,ball2006robust,sevilla2015effect}, 
Support Vector Machines~\citep{Fadely2012}, and
classifier combination strategies~\citep{kim2015hybrid}.

Almost all star-galaxy classifiers published in the literature use
the reduced summary information available from astronomical catalogs.
Constructing catalogs requires careful engineering and considerable domain expertise
to transform the reduced, calibrated pixel values that comprise an image into suitable features,
such as magnitudes or shape information of an object.
In a branch of machine learning called \textit{deep learning},
features are not designed by human experts;
they are learned directly from data by deep neural networks.
Deep learning has been applied successfully to galaxy morphological
classification in Sloan Digital Sky Survey~\citep[SDSS;][]{dieleman2015rotation}
and Cosmic Assembly Near-infrared Deep Extragalactic Legacy
Survey~\citep[CANDELS;][]{huertas2015catalog} and to photometric redshift
estimation~\citep{hoyle2015measuring},
but it has not yet been applied to the problem of source classification.

\RC{FYI, a raw pixel value is observed directly from the telescope and is generally worthless. It must be reduced (correcting for the optics of the telescope and observational imprints) and calibrated (so that photon counts mean a physical value) before being useful in any sense.}

In this paper, we present a star-galaxy classification framework
that uses a convolutional neural network (ConvNet) model directly on the images
from the SDSS and the Canada-France-Hawaii Telescope Lensing Survey (CFHTLenS).
We compare its performance with a standard machine learning technique
that uses the reduced summary information from catalogs,
and we demonstrate that our ConvNet model is able to produce accurate and
well-calibrated probabilistic classifications with very little feature
engineering by hand.
In Section~\ref{sec:data}, we describe the data sets used in this paper
and the pre-processing steps for preparing the image data sets.
We provide a brief overview of deep learning and ConvNets
in Section~\ref{sec:deep_learning}, and
discuss our strategy for preventing overfitting in Section~\ref{sec:overfitting}.
In Section~\ref{sec:tpc}, we describe a state-of-the-art tree-based
machine learning algorithm, to which the performance of our ConvNet model is
compared.
We present the main results of our ConvNet model in
Section~\ref{sec:results_and_discussion}, and
we outline our conclusions in Section~\ref{sec:conclusions}.

@@

\begin{figure}
  \centering
  \includegraphics[width=\columnwidth]{figures/sdss_color_color.pdf}
  \caption{
    Hexbin plots showing the color-color space distribution of galaxies (blue, top row) and stars (green, bottom row)
    in the SDSS data set used in this analysis.
    It is easy to see that there is significant overlap between stars
    and galaxies in these color-color spaces. 
  }
  \label{fig:sdss_color_color}
\end{figure}

\RC{It might be helpful to have a colorbar with these plots to show relative densities. I also wonder if the extra whitespace might be reduced somehow. Finally, I actually think a ratio plot might be best her, where the hexbin shows the ratio of the normalized counts.}

The Sloan Digital Sky Survey~\citep[SDSS;][]{york2000sloan}
phases I--III obtained photometric data in five bands,
$u$, $g$, $r$, $i$, and $z$,
covering 14,555 square degrees, more than one-third of the entire sky.
The resulting catalog contains photometry of over 300 million stars and galaxies
with a limiting magnitude of $r \approx 22$,
making the SDSS one of the largest sky surveys ever undertaken.
The SDSS also conducted an expansive spectroscopic follow-up of
more than three million stars and galaxies~\citep{eisenstein2011sdss}.
In this paper, we use a subset of the photometric and spectroscopic data
contained within the Data Release 12~\citep[DR12;][]{alam2015eleventh},
which is publicly available through the online CasJobs
server\footnote{http://skyserver.sdss.org/casjobs/}~\citep{li2008casjobs}.

Using the CasJobs server, we randomly select a total of 65,000 sources, which are either stars or galaxies.
We exclude some bad photometric observations as follows.
We consider only objects whose $r$-band magnitude is less than 40;
there are no warning flags in the spectroscopic measurement (\texttt{zWarning = 0});
the half-light radius in the $r$ band is less than 30 arc seconds
as measured by the exponential and de Vaucouleurs light profiles;
the error on the spectroscopic redshift measurement is less than 0.1; and
the spectroscopic redshift is less than 2.

To create training images,
we obtain the image FITS files for SDSS fields containing these objects
in five photometric bands: $u$, $g$, $r$, $i$, and $z$.
We use the astrometry information in the FITS headers
in the \texttt{Montage}\footnote{http://montage.ipac.caltech.edu/} software
to align each image to the reference ($r$-band) image.
We then use \texttt{SExtractor} to find the pixel positions of
the $65\,000$ objects we have selected,
and to center each object in the cutout image.
We also convert the flux values to inverse hyperbolic sine magnitudes
\citep[also known as luptitudes;][]{lupton1999modified}.
Finally, in order to account for the effect of Galactic dust,
extinction corrections in magnitudes are applied
following \cite{schlegel1998maps}.
In the end, we have cutout images of size $48\times48$ pixels
with luptitude values in each pixel.
We note that we have experimented with increasing the pixel dimensions to
$60\times60$ and $72\times72$ pixels, but do not find noticeable
improvement in the performance of our model.

In the end, we have 17,344 stars and 47,656 galaxies available
for the training and testing processes.
The apparent magnitudes range from $10.7 < r < 23.1$,
and the galaxies in this sample have a mean redshift of $z \sim 0.36$.
The $ugriz$ color-color space of the sample is plotted in
Figure~\ref{fig:sdss_color_color}.
We randomly split the objects into training, held-out validation,
and blind test sets of size 40,000, 10,000, and 15,000, respectively.
We note that cross-validation is often avoided in deep learning
in favor of hold-out validation,
since cross-validation is computationally expensive.
We also note that we perform a blind test, and the test set is not used in any
way to train or calibrate the algorithms.

\RC{Be useful to state what the ratio of Stars to galaxies are in these three data sets (train, validate, test). Also, a referee will likely ask how they are distributed in magnitude and color.}

@@

\begin{figure}
  \centering
  \includegraphics[width=\columnwidth]{figures/clens_color_color.pdf}
  \caption{
    Hexbin plots showing the color-color space distribution of stars and
    galaxies in the CFHTLenS data set used in this analysis.
  }
  \label{fig:clens_color_color}
\end{figure}

We also use photometric data from
the Canada-France-Hawaii Telescope Lensing Survey
\cite[CFHTLenS\footnote{http://www.cfhtlens.org/};][]
{heymans2012cfhtlens,erben2013cfhtlens,hildebrandt2012cfhtlens}.
This catalog consists of more than twenty five million objects
with a limiting magnitude of $i_{\text{AB}} \approx 25.5$. 
It covers a total of 154 square degrees
in the four fields (named W1, W2, W3, and W4)
of the CFHT Legacy Survey~\citep[CFHTLS;][]{gwyn2012canada}
observed in the five photometric bands:
$u$, $g$, $r$, $i$, and $z$.

@@

We obtain FITS images for each 1 square degree CFHTLenS pointing
that contains objects with spectroscopic labels.
We create cutout images of size $96\times96$ pixels by using a similar method
to that described in Section~\ref{sec:sdss}.
Finally, images are downscaled to $48\times48$ pixels
to reduce the computational cost.

In the end, we have 8,545 stars and 57,843 galaxies available
for the training and testing processes.
The apparent magnitudes range from $13.9 < r < 25.6$,
and the galaxies in this sample have a mean redshift of $z \sim 0.59$.
The $ugriz$ color-color space of the sample is plotted in
Figure~\ref{fig:clens_color_color}.
We randomly split the objects into training, held-out validation, and
blind test sets of size $40\,000$, $10\,000$, and $13\,278$, respectively.

\RC{ Again the S:G ratio in the data sets and mag and color distributions. Referee will likely ask, so you can put something in here to simplify review process.}
\section{Deep Learning}
  \label{sec:deep_learning}

@@

The sigmoid function, $\sigma(x)=1/\left(1+e^{-x}\right)$,
has frequently been used as an activation function, but it has recently fallen out of favor,
and the hyperbolic tangent function, $\sigma(x)=\tanh(x)$, is generally used instead.
The most popular non-linearity at present, however, is the rectified linear unit
\citep[ReLU;][]{nair2010rectified}, $\sigma(x)=\max(0, x)$.
ReLUs generally allow much faster
training of deep neural networks with many layers.
However, ReLU units can sometimes result in dead neurons whose output is always zero.
To mitigate this problem, we use leaky ReLUs that have a small, non-zero slope
in the negative region,
\begin{equation}
  \sigma(x) =
    \begin{cases}
      x & \mbox{if } x \geq 0 \\
      0.01x & \mbox{if } x < 0.
    \end{cases}
\end{equation}

Many deep learning models use feedforward neural network architectures with multiple layers,
where each neuron in one layer is connected to the neurons of the subsequent layer.
A schematic representation is shown in Figure~\ref{fig:neuron_b}.
All layers except the input and output layers are conveniently called hidden layers.
With multiple hidden layers, deep neural networks can model complex nonlinear
relationships, and they have had great success in multimedia tasks where humans traditionally
outperformed computers, such as speech recognition, language processing, and image classification.

@@

\RC{You do not talk anywhere on the software you actually use to do this work. You should introduce Python, Lasagna, Theano, and any other relevant software. You also should cite them. Finally, what hardware was used for these tests?}

@@

\begin{table*}
  \centering
  \caption{Summary of ConvNet architecture and hyperparameters. Note that pooling layers have no learnable parameters. \RC{ This might be much easier to understand if it is a flowchart, like we did in INFO490 with Keras. Is that possible? I say this as it would better  match your toy ANN in  a previous figure.}}
  \label{table:hyperparamters}
  \begin{tabular}{ccccccc}
    \hline
    type            & filters & filter size & padding & non-linearity & initial weights & initial biases \\
    \hline
    convolutional   & 32          & $5\times5$  & -       & leaky ReLU    & orthogonal      & 0.1            \\
    convolutional   & 32          & $3\times3$  & 1       & leaky ReLU    & orthogonal      & 0.1            \\
    pooling         & -           & $2\times2$  & -       & -             & -               & -              \\
    convolutional   & 64          & $3\times3$  & 1       & leaky ReLU    & orthogonal      & 0.1            \\
    convolutional   & 64          & $3\times3$  & 1       & leaky ReLU    & orthogonal      & 0.1            \\
    convolutional   & 64          & $3\times3$  & 1       & leaky ReLU    & orthogonal      & 0.1            \\
    pooling         & -           & $2\times2$  & -       & -             & -               & -              \\
    convolutional   & 128         & $3\times3$  & 1       & leaky ReLU    & orthogonal      & 0.1            \\
    convolutional   & 128         & $3\times3$  & 1       & leaky ReLU    & orthogonal      & 0.1            \\
    convolutional   & 128         & $3\times3$  & 1       & leaky ReLU    & orthogonal      & 0.1            \\
    pooling         & -           & $2\times2$  & -       & -             & -               & -              \\
    fully-connected & 2048        & -           & -       & leaky ReLU    & orthogonal      & 0.01            \\
    fully-connected & 2048        & -           & -       & leaky ReLU    & orthogonal      & 0.01            \\
    fully-connected & 2           & -           & -       & softmax       & orthogonal      & 0.01            \\
    \hline
  \end{tabular}
\end{table*}

@@

\RC{ I wonder why? Is this somehow relevant?}

We have performed a manual search to explore more than 200 combinations of different architectures
and hyperparameters to find an architecture that minimizes the loss function
(Equation~\ref{eq:cross_entropy_all}) on the validation set of the SDSS data.
The architecture described in this section provides the best performance on
the SDSS validation set.
To test how this model performs across different, related data,
we use the same architecture on the CFHTLenS data set.

@@

We note that training our network takes about forty hours on an NVIDIA Tesla K40 GPU.

@@

Our convolution neural network has $11\times10^6$ learnable parameters,
while there are only $4\times10^4$ images in the training set.
As a result, the model is very likely to \emph{overfit} without regularization.
In this section, we describe the techniques we used to minimize overfitting.

@@

We continue the sampling process until
we reach a predefined number of combinations,
and we finally use Equation~\ref{eq:posterior_bmc} to compute
the posterior probability that a source is a star (or a galaxy).
In this paper, we use a $q$ value of three,
and we consider a total of $1\,000$ model combinations.

@@

TPC uses classification trees,
a type of prediction tree that is designed to
provide a classification or predict a discrete category.
Prediction trees are built by asking a sequence of questions
that recursively split the data into branches
until a terminal leaf is created
that meets a stopping criterion
(\eg a minimum leaf size).
The optimal split dimension is decided by
choosing the attribute that maximizes
the \textit{Information Gain} ($I_G$), which is defined as

@@

\noindent
where $f_g$ is the fraction of galaxies in the training data.
At each node in our tree,
we scan all dimensions to identify the split point that
maximizes the information gain as defined by Equation~\ref{eq:information_gain},
and select the attribute that maximizes the overall impurity index.

@@

We train two TPC models on the SDSS data set by using different sets of
attributes.
The first model, which we denote $\rmn{TPC}_{\rmn{phot}}$, is trained with
only nine photometric attributes:
the extinction-corrected model magnitudes in five bands
($u$, $g$, $r$, $i$, $z$)
and their corresponding colors
($u-g$, $g-r$, $r-i$, $i-z$).
The second model, which we denote $\rmn{TPC}_{\rmn{morph}}$, is trained with
the concentration parameter in each band in addition to the magnitudes and
colors, for a total of fifteen dimensions.
The concentration is defined as the difference between the cmodel
and PSF magnitude: $\rmn{concentration}\equiv\rmn{psfMag}-\rmn{cModelMag}$.
The SDSS pipeline uses a parametric method based on the concentration,
an object is classified as a galaxy if $\rmn{concentration}>0.145$.
We find that the concentration is an excellent morphological feature
for star-galaxy separation, and including more morphological features
does not show noticeable improvement in performance.
The concentration is a good example of carefully handcrafted feature extraction;
we show in Section~\ref{sec:results_and_discussion}
that ConvNets do not require such feature engineering.

@@

Probabilistic classification models can be considered as
functions that output a probability estimate of each source
to be in one of the classes (\eg a star or a galaxy).
The simplest way to use the probability output is to
convert it into a class label by using a threshold (a probability cut).
Although many studies often choose the threshold by setting it to a fixed value,
such as $P_{\rmn{cut}} = 0.5$ \citep[\eg][]{henrion2011bayesian, Fadely2012},
choosing $0.5$ as a threshold is not the best choice
for an unbalanced data set, where galaxies outnumber stars.
Furthermore, setting a fixed threshold ignores the operating condition 
(\eg science requirements, stellar distribution, misclassification costs)
where the model will be applied.

@@

We also use the mean squared error
(MSE; also known as the Brier score~\citep{brier1950verification})
as a performance metric. We define MSE as
\begin{equation} \label{eq:mse}
  \rmn{MSE} = \frac{1}{N} \sum^{N}_{j=1}
  \left( y_j - \hat{y}_j \right)^2,
\end{equation}
The MSE can be considered as both a score function that quantifies
how well a set of probabilistic predictions is calibrated,
or a loss function.

@@

\RC{Why 1000 here?}

The problem with a binning approach is either too few or too many bins
can distort the evaluation of calibration performance.
Thus, we adopt a calibration measure based on
overlapping binning~\citep{caruana2004data}.
We order the predicted values $P_\rmn{class}$ and
put the first 1,000 elements in the first bin.
We calculate the true probability $P_\rmn{gal}$ by counting the true galaxies
in this bin.
The calibration error for this bin is $\mid P_\rmn{gal} - P_\rmn{class} \mid$.
We then repeat this for the second bin (2 to 1,001),
the third bin (3 to 1,002), and so on, and average the binned calibration errors.
Thus, the overall calibration error is given by
\begin{equation}
  \rmn{CAL} = \frac{1}{N-s} \sum_{b=1}^{N-s} \sum_{j=b}^{b+s-1}
  \bigg| P_{\rmn{class},j} - \frac{\sum_{j=b}^{b+s-1} P_{\rmn{gal},j}}{s} \bigg|,
\end{equation}
where $s=1\,000$ is the bin length.
